### 2018-01-11 includes/abstract.tex
Writing sections in paper.

    In this paper we present StringFuzz: an SMT-LIB problem fuzzer and generator for string solvers. With it we analyse \numSolvers{} mature string solvers: \theSolvers{}. We use it to craft inputs that elicit poor performance in Z3str3 and CVC4, and provide analyses of the causes of the performance degradations. Our experimental data suggest that there are classes of problems that will always be hard for Z3str3 and easy for CVC4, and vice versa, due to the limitations of the core algorithms implemented by those solvers. We also present several minor bugs in Z3str3 that were exposed by StringFuzz. Finally, we provide a rich repository of problems in SMT-LIB format, generated both by StringFuzz and by hand, to encourage other solver authors to likewise improve their solvers.

>>>>

    In this paper we present \fuzzer{}: an \smt{} problem fuzzer and generator for string solvers. With it we analyse \numSolvers{} mature string solvers: \theSolvers{}. We use it to craft inputs that elicit poor performance in \us{} and \cvc{}, and provide analyses of the causes of the performance degradations. Our experimental data suggest that there are classes of problems that will always be hard for \us{} and easy for \cvc{}, and vice versa, due to the limitations of the core algorithms implemented by those solvers. We also present several minor bugs in \us{} that were exposed by \fuzzer{}. Finally, we provide a rich repository of problems in \smt{} format, generated both by \fuzzer{} and by hand, to encourage other solver authors to likewise improve their solvers.

### 2018-01-11 includes/analysis.tex
Added section about partial order eqc comparison solution.
Writing sections in paper.

    Here we present instances we generated, and our empirical results.

@@

        Here we describe the problem. CVC4 did poorly. Z3str3 did well.

@@

        Here we describe the problem. CVC4 did poorly. Z3str3 did well.

@@

        Here we describe the problem. CVC4 did poorly. Z3str3 did well.

@@

        Here we describe the problem. CVC4 did well. Z3str3 did poorly.

    Here we talk about Lazy-Hard and Eager-Hard, and how the algorithms of CVC4 and Z3str3 differ, eliciting the observed behaviour.

        Here is why CVC4 does poorly and Z3str3 does well.

        Here is why CVC4 does well and Z3str3 does poorly.

>>>>

    Here we present two particular sets of problems generated by \fuzzer{}: Eager-Hard and Lazy-Hard. These two problem sets were unique in their ability to elicit poor performance in one solver, but not another. Full details of test scripts, configurations, and problems that were used in the experiment can be found in the Appendix.

    \textbf{TODO:}

    \begin{enumerate}
        \item supply machine environment description
        \item supply install scripts that produce trace builds
        \item supply test scripts that invoke the solvers and produce traces
    \end{enumerate}

\section{Results}

    \subsection{Performance on Lazy-Hard}

        On this problem set \cvc{} did poorly, but \us{} did well. The graphs in figure X show the performance results for both solvers.

        \cvc{} showed the following behaviour in its log file:

        \begin{enumerate}
            \item Loads all facts
            \item Flattens concats
            \item Computes equivalence classes
            \item Discovers conflict early
                {\scriptsize\begin{verbatim}
CONFLICT: Eq engine conflict : (and (= "woof" (str.++ "woof" lsym_5 lsym_5 lsym_5 lsym_5
lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5)) (= "" lsym_5)
(= "" var3) (= "" var5) (= "" var7) (= "" var9) (= "" var11) (= "" var13) (= "" var15)
(= "" var17) (= "" var19) (= "" var21) (= "" var23) (= "" var25) (= "" var27) (= "" var29)
(= "" var30) (= (str.++ "woof" var3 var5 var7 var9 var11 var13 var15 var17 var19 var21
var23 var25 var27 var29 var30) (str.++ "meow" var43 var45 var47 var49 var51 var53 var55
var57 var59 var61 var63 var65 var67 var69 var70)) (= "" var43) (= "" var45) (= "" var47)
(= "" var49) (= "" var51) (= "" var53) (= "" var55) (= "" var57) (= "" var59) (= "" var61)
(= "" var63) (= "" var65) (= "" var67) (= "" var69) (= "" var70) (= "meow" (str.++ "meow"
lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5 lsym_5
lsym_5 lsym_5)))
                \end{verbatim}}
            \item Gets caught doing some sort of check
            \item Infers inequality
        \end{enumerate}

        \us{} showed the following behaviour in its log file:

        \begin{enumerate}
            \item loads all facts
            \item immediately spots inconsistency:
                {\scriptsize\begin{verbatim}
-------- [str] new_eq_check ../src/smt/theory_str.cpp:1864 ---------
checking whether (str.++ "woof" var2) and (str.++ "meow" var42) can be equal
------------------------------------------------
-------- [str] new_eq_check ../src/smt/theory_str.cpp:1867 ---------
inconsistency detected: (str.++ "woof" var2) cannot be equal to (str.++ "meow" var42)
------------------------------------------------
-------- [str] assert_axiom ../src/smt/theory_str.cpp:179 ---------
asserting (not (= (str.++ "woof" var2) (str.++ "meow" var42)))
------------------------------------------------
                \end{verbatim}}
        \end{enumerate}

        \textbf{TODO:}

        \begin{enumerate}
            \item \us{} cactus plot
            \item \cvc{} cactus plot
            \item comparison plot
        \end{enumerate}

        On this problem set \cvc{} did well, but \us{} did poorly. The graphs in figure X show the performance results for both solvers.

        \cvc{} showed the following behaviour in its log file:

        \begin{enumerate}
            \item Loads all facts
            \item Flattens concats
            \item Computes equivalence classes
            \item Finds solution
        \end{enumerate}

        \us{} showed the following behaviour in its log file:

        \begin{enumerate}
            \item Loads all facts
            \item Tries assigning solution by "dragging" it across the variables
            \item Creates a ton of equivalence classes
            \item Gets caught comparing equivalence classes
            \item Finds solution
        \end{enumerate}

        \textbf{TODO:}

        \begin{enumerate}
            \item \us{} cactus plot
            \item \cvc{} cactus plot
            \item comparison plot
        \end{enumerate}

@@

        No data on this yet.

@@

        No data on this yet.

    This section presents our analysis of the observed behaviour of \cvc{} and \us{} on Lazy-Hard and Eager-Hard problems.

@@

        Here is why \cvc{} does poorly and \us{} does well.

@@

        Here is why \cvc{} does well and \us{} does poorly.

\section{\us{} Recommendation}

    This section outlines a change for \us{}, which we believe will improve its performance on the Eager-Hard problem set and similar problems. We recommend a partial order treatment of equivalence class comparisons. First, we describe a partial order of equivalence classes. Then, we use it in equivalence class comparisons and more closely approximate equality between them.

    \textbf{TODO:}

    \begin{enumerate}
        \item definition of partial order
        \item definition of current class relation
        \item definition of class relation under partial order
    \end{enumerate}

### 2018-01-11 includes/introduction.tex
Writing sections in paper.

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and practically useful.

    Moreover a standard input language, SMT-LIB, has been developed, and many mature solvers support it.

>>>>

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and practically useful. Moreover a standard input language, \smt{} (versions 2, 2.1, and 2.5), has been developed and many mature solvers support it.

    \subsection{Prior Work}

        Other fuzzers exist, but none of them support the theory of strings. We found these:

        \begin{itemize}
            \item http://fmv.jku.at/ddsmt/
            \item http://fmv.jku.at/fuzzsmtbv/
            \item http://fmv.jku.at/fuzzsmt/
            \item http://fmv.jku.at/deltasmt/
            \item http://fmv.jku.at/votesmt/
            \item http://www.verit-solver.org/papers/smt2015.pdf
            \item http://smt2015.csl.sri.com/wp-content/uploads/2015/07/smt2015-slides-deharbe.pdf
        \end{itemize}

    \subsection{Contributions}

        In this paper, we present the following contributions:

        \begin{itemize}
            \item StringFuzzG, a generator of \smt{} string problems
            \item StringFuzzX, a transformer of \smt{} string problems
            \item Experiments that show evidence of poor performance in \us{} and \cvc{}
            \item A recommended fix to \us{} that alleviates the poor performance
            \item A suite of \smt{} problems
        \end{itemize}

        In the following sections, we elaborate on each of these items.

### 2018-01-11 includes/problems.tex
Writing sections in paper.

    \subsection{StringFuzz-Generated}

        Here we describe the StringFuzz-generated suite.

    \subsection{StringFuzz-Transformed}

        Here we describe the StringFuzz-transformed suite.

@@

        Here we describe the unit test suite.

>>>>

    Along with this paper we present several suites of ready-made problems produced by StringFuzz. These suites can be found on \problemRepo{}.

    \subsection{\fuzzer{}-Generated}

        This suite contains the following problem sets:

        \begin{itemize}
            \item Lengths
            \item Lengths-Concats
            \item Concats
            \item Concats-Extracts
            \item Overlaps
            \item Regex
            \item Many-Regexes
            \item Regex-Deep
            \item Regex-Pair
            \item Regex-Lengths
            \item Random??
        \end{itemize}

        Most of these problem sets either have a dedicated generator in \generator{}, or are variations of one. While it is likely that some of the pre-made suites will be immediately useful in analysing a solver, we also found that mass randomisation was helpful in finding some bugs, and we encourage this pattern as well. \generator{} is designed to be used in an exploratory fashion, and the following usage has helped us in our own experiments:

\begin{verbatim}
while true; do stringfuzzg --random regex | z3str3 -in; done
\end{verbatim}

    \subsection{\fuzzer{}-Transformed}

        This suite contains the following problem sets:

        \begin{itemize}
            \item Amazon??
            \item Regex??
        \end{itemize}

        Most of these problem sets were generated by using one root problem and transforming it with \transformer{} into many similar problems.

@@

        This suite covers corner cases of solver behaviour. We have used the instances in this suite to find bugs in \us{}.

    \subsection{Results}

        This section describes the bugs in \us{} that were discovered (and subsequently fixed) by testing it with the \fuzzer{} problem suites.

        \begin{enumerate}
            \item length lookup using eqc root
            \item length-testing bug
        \end{enumerate}

### 2018-01-11 includes/stringfuzz.tex
Writing sections in paper.

\section{StringFuzz}

    Here we present StringFuzz.

    \subsection{Generators}

        Here we describe our generators (concats, overlaps, etc.).

    \subsection{Transformers}

        Here we describe our transformers (graft, fuzz, etc.).

@@

        Here we describe our utilities (stringstats).

>>>>

\section{\fuzzer{}}

    This section presents \fuzzer{}. \fuzzer{} is implemented as a Python package, and comes with several executables to generate, transform, and analyse \smt{} string problems. Its source code repository is available at \sourceRepo{}. It can either be installed from source, or from the Python PIP package repository.

    \subsection{Generators: \generator{}}

        \generator{} is available as a command-line tool, \texttt{stringfuzzg}. It supports several generators and options that control its output. The generators it implements are:

        \begin{itemize}
            \item concats
            \item lengths
            \item overlaps
            \item regex
            \item random-text
        \end{itemize}

    \subsection{Transformers: \transformer{}}

        \transformer is available as a command-line tool, \texttt{stringfuzzx}. It supports several transformers and options that control its output and input. The transformers it implements are:

        \begin{itemize}
            \item concats
            \item lengths
            \item overlaps
            \item regex
            \item random-text
        \end{itemize}

@@

        Some utilities are also provided with \fuzzer{}. They are: \texttt{stringstats}, \texttt{stringmerge}, \texttt{smtparse}, and \texttt{smtscan}.

        \begin{description}

            \item[\texttt{stringstats}] \hfill \\ \\
                This tool takes as input an \smt{} problem, and outputs its properties. The properties are:

                \begin{itemize}
                    \item number of string variables
                    \item number of string literals
                    \item max/median syntactic depth of expressions
                    \item max/median length of literals
                \end{itemize}

                \hfill

            \item[\texttt{stringmerge}] \hfill \\ \\
                This tool takes as input two \smt{} problems and combines them into one problem.

        \end{description}

    \subsection{Architecture}

        Should I describe its architecture?
